---
title: "Batch Processing for my AST data"
author: "Ashma"
date: "April 7, 2020"
output:
  prettydoc::html_pretty
---

#### Repeat the exercise from the Batch Processing Lecture (7th April), but do it using real data sets rather than purely simulated. Check with folks in your lab to see if there are multiple data sets available for analysis. Stick to simple data analyses and graphics, but try to set it up as a batch process that will work on multiple files and save summary results to a common file.

```{r}
# -----------------------------------------
# FUNCTION reg_stat
# description: fits linear model by extracting statistics and summary
# inputs: two column dataframe x and y
# outputs: slope,p-value and r2
###########################################
reg_stat <- function(d=NULL) {
             if(is.null(d)) {
               xVar <- "Cows"
               yVar <- "Humans"
               d <- data.frame(xVar,yVar)
             }
  . <- lm(data=d,d[,2]~d[,1])
  . <- summary(.)
  stat_list <- list(slope=.$coefficients[2,1],
                    pVal=.$coefficients[2,4],
                    r2=.$r.squared)

return(stat_list)

} # end of reg_stat
#-------------------------------------------
```


```{r}
library(TeachingDemos)
char2seed("Lockdown Assignment")

#--------------------------------------------
# Global variables
file_folder <- "AST_data/"
n_files <- 18
file_out <- "AST_stat_summary.csv"
#--------------------------------------------
drug_list <- list.files(path=file_folder)

# Create data frame to hold file summary statistics
ID <- seq_along(drug_list)
fileName <- drug_list
slope <- rep(NA,n_files)
pVal <- rep(NA,n_files)
r2 <- rep(NA,n_files)
stat_output <- data.frame(ID,fileName,slope,pVal,r2)
```


```{r}
# batch process by looping through individual files
for (i in seq_along(drug_list)) {
  AST <- read.table(file=paste(file_folder,drug_list[i],sep=""),
                     sep=",",
                     header=TRUE) # read in next data file

  AST_clean <- AST[complete.cases(AST),] # removing missing data
  . <- reg_stat(AST) # pull regression stats from clean file
  stat_output[i,3:5] <- unlist(.) # unlist, copy into last 3 columns
  
}
# set up output file and incorporate time stamp and minimal metadata
  write.table(cat("# Summary stats for ",
                    "batch processing of regression models","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# AC","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=file_out,
                    row.names="",
                    col.names="",
                    sep=""))
  
# now add the data frame
  write.table(x=stat_output,
              file=file_out,
              row.names=FALSE,
              col.names=TRUE,
              sep=",",
              append=TRUE)
```