---
title: "Structured Programming"
author: "Ashma"
date: "March 25, 2020"
output: html_document
---

### Organizing code with Structured Programming

```{r}
# load library -----------------------------
library(ggplot2)

# create global variables -----------------------------

vari <- data.frame(n_name=c("cows","humans"),# 2 groups
                n_size=c(25,25), # 25 in each group
                n_mean=c(50,21), # means of groups
                n_sd=c(10,7))

  

# -------------------------------
# FUNCTION get_data
# description: generate data frame in correct orientation
# inputs: numberic variables normally distributed, in vector form 
# outputs: data frame
##################################

get_data <- function(vari) { 
  amr <- c(rnorm(n=vari$n_size[1],mean=vari$n_mean[1],sd=vari$n_sd[1]),
                 rnorm(n=vari$n_size[2],mean=vari$n_mean[2],sd=vari$n_sd[2]))
  spp <- rep(vari$n_name,vari$n_size) 
id <- 1:(sum(vari$n_size))
    df <- data.frame(id,spp,amr)

return(df)

} # end of get_data

#---------------------------------


# -------------------------------
# FUNCTION summarize_ANOVA
# description: run ANOVA model on data
# inputs: normally distributed variables of numeric
# outputs: ANOVA model
##################################
summarize_ANOVA <- function(df) {
  z <- aov(amr ~ spp, data=df)
ano_sum <- summary(z)

return(ano_sum)

} # end of calculate_ANOVA
#--------------------------------


# -------------------------------
# FUNCTION datPlot
# description: plotting data
# inputs: data frame with variables of numeric
# outputs: box plot
##################################
datPlot <- function(df) {
ano_plot <- ggplot(data=df,
                 aes(x=spp,
                     y=amr,
                     fill=spp)) +
  geom_boxplot()
 
print(ano_plot)
message("Message: BoxPlot created for ANOVA!")

} # end of graph_normal
#---------------------------------

```


### Using the functions

```{r}
df <- get_data(vari) # temporary variable 
summarize_ANOVA(df) 
datPlot(df)
```


### Modifying the function by adding linear regression

```{r}
# -------------------------------
# FUNCTION: pos_reg
# description: calculate regression
# inputs: data frame in which x is categorical variable and y is continuous variable
# outputs: poisson regression
##################################
pos_reg <- function(df) {
posModel <- glm(df[,1]~df[,2], data=df, family=poisson())

return(summary(posModel))

} # end of pos_reg
#---------------------------------


#---------------------------------
###################################
# FUNCTION: posPlot 
# description: plot a poisson regression
# INPUTS: dataframe in which the first col is the x variable and second is the y
# OUTPUTS: a plot of the regression
#-------------------------
posPlot <- function(df) {
  
  posPlot <- ggplot(data=df,aes(x=spp,y=amr)) +
           geom_point() +
           geom_smooth(method="glm",se=TRUE)  
print(posPlot)


} # end of posPlot
#---------------------------------

```

### Adding the new functions to the previous ones

```{r}
df <- get_data(vari) # temporary variable 
summarize_ANOVA(df) 
datPlot(df)
pos_reg(df)
posPlot(df)
```

### Optional. If time permits and you have the skills, try putting your program inside of a for loop and repeat the analysis with a different stochastic data set (each time you call a function that invokes the random number generator, it will create a new set of data for you to process). Can you create a data structure to store the summary statistics created in each pass through the loop? If not, your program will work, but it will only show the results from the final replicate (the previous results will be written over each time you traverse the loop).

```{r}
nreps <- 20
p_val <- matrix(NA, nrow=nreps, ncol=1)

for (i in 1:nreps) {
  df <- get_data(vari)
  aov <- summarize_ANOVA(df)
  av <- unlist(aov)
  p_val[i,] <- av[9]
}

print(p_val)
```