---
title: "Matrices, Lists and DataFrames"
author: "Ashma"
date: "February 12, 2020"
output: html_document
---


## Assign to the variable n_dims a single random integer between 3 and 10.

```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```

### Create a vector of consecutive integers from 1 to n_dims2.

```{r}
vec <- seq(1:(n_dims^2))
```

### Use the sample function to randomly reshuffle these values.

```{r}
fx <- sample(vec)
```

### create a square matrix with these elements.

```{r}
sq_m <- matrix(data = fx, nrow = sqrt(length(fx))) # using square root of the length of vector for nrow
```

### print out the matrix

```{r}
print(sq_m)
```

### find a function in r to transpose the matrix and print it out again.

```{r}
sq_mt <- t(sq_m) # tranpose the matrix
print(sq_mt) 
```

### calculate the sum and the mean of the elements in the first row and the last row.

```{r}
sum(sq_mt[1,]) # calculate sum of elements in first row
sum(sq_mt[nrow(sq_mt), ]) # calculate sum of elements in last row
mean(sq_mt[1,]) # calculate mean of elements in first row
mean(sq_mt[nrow(sq_mt), ]) # calculate the mean of elements in last row
```

### read about the eigen() function and use it on your matrix.

```{r}
eg<- eigen(sq_mt) # calculates eigenvalues and eigenvectors of matrix
```

### look carefully at the elements of $values and $vectors. What kind of numbers are these?

```{r}
print(eg)
```

### dig in with the typeof() function to figure out their type.

```{r}
typeof(eg$values) # shows what is the type of elements within the variable
typeof(eg$vectors)
```

## Create a list with the following named elements:

### my_matrix, which is a 4 x 4 matrix filled with random uniform values

```{r}
my_matrix <- matrix(data = runif(16), nrow=4)
print(my_matrix)
```

### my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.

```{r}
my_logical <- (rnorm(100) > 1) # create a logical vector using inequality
print(my_logical)
```

### my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
my_letters <- sample(letters[1:26]) # creates vector of random order lower case letters.
print(my_letters)
```

### create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
my_list <- list(my_matrix[2,2], my_logical[2], my_letters[2]) # create a list 
print(my_list)
```

### use the typeof() function to confirm the underlying data types of each component in this list

```{r}
typeof(my_list[[1]]) # using [[]] to capture element of the list
typeof(my_list[[2]])
typeof(my_list[[3]])
```

### combine the underlying elements from the new list into a single atomic vector with the c() function.

```{r}
new_vec <- c(my_list[[1]],my_list[[2]],my_list[[3]]) # concatanate into a vector
print(new_vec)
```

### what is the data type of this vector?

```{r}
typeof(new_vec) # data type is charcater
```

## Create a data frame with two variables (= columns) and 26 cases (= rows)

### call the first variable my_unis and fill it with a random uniform value from 0 to 10
### call the second variable my_letters and fill it with 26 capital letters in random order.

```{r}
set.seed(1234)
my_unis <- runif(n=26, min=0, max=10) # create variable with random unifrom values from 0 to 10
my_letters <- sample(LETTERS[1:26]) # create another variable with 26 random order capital letters
my_dframe <- data.frame(my_unis,my_letters,stringsAsFactors = FALSE) # create dataframe
str(my_dframe) # view structure of dataframe
```

### for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA

```{r}
my_dframe$my_unis[sample(my_dframe$my_unis, size= 4,replace = FALSE)] <- NA # replacing values in 4 rows of my_unis with NA
head(my_dframe$my_unis) # view the data
```

### for the first variable, write a single line of R code to identify which rows have the missing values.

```{r}
which(is.na(my_dframe$my_unis)) # finds rows which are NAs
```

### for the second variable, sort it in alphabetical order

```{r}
my_dframe$my_letters <- my_dframe$my_letters[order(my_dframe$my_letters)] # sorting in alphabetical order
str(my_dframe)
```

### calculate the column mean for the first variable

```{r}
mean(my_dframe$my_unis, na.rm = TRUE) # mean of first variable with NA
```