---
title: "Vectors In R"
author: "Ashma"
date: "February 5, 2020"
output: html_document
---

---

# Assignment operator

### _Assign each expression to the value of the variable z and print the value stored in z_

```{r}
x <- 1.1 # assigning values 
a <- 2.2
b <- 3.3

z <- x^a^b # assigning the expression to z
print(z) # printing the result for z

```

```{r}
z <- x^a*b
print(z)

```

``` {r}
z <- 3*x^3 + 2*x^2 + 1
print(z)

```

# Create Vectors

### _Using the rep and seq functions, create the following vectors_

```{r}
seq_A <- seq(from=1, to=8, by=1) # creates an ascending sequence from 1 to 8 
seq_D <- seq(from=7, to=1, by=-1) # creates a descending sequence from 7 to 1
vec1 <- c(seq_A, seq_D) # concatanate the 2 sequences and create a vector
print(vec1)

```

```{r}
seq_R <- seq(from=1, to=5, by=1)
vec2 <- rep(x=seq_R, times=seq_R) # repeats each element in the vector as its own value
print(vec2)

```

```{r}
seq_I <- seq(from=5, to=1, by=-1)
vec3 <- rep(x=seq_I, c(seq(1,5))) # repeats element in vector by sequence as first element once and last element 5 times
print(vec3)

```

# Polar Coordinates with random uniform numbers

### _Using one of R's inverse trigonometry functions (asin(), acos(), or atan()), convert random uniform numbers into polar coordinates_

```{r}
rnum <- runif(2) # creating two random numbers 
p <- rnum[1] # assigning first random number as p
q <-rnum[2] # assigning second random number as q
pol <- sqrt(p^2+q^2) # calculate value of polar coordinate r
theta <- atan(q/p) # calculate angular coordinate theta
print(pol)
print(theta)

```

# Playing with vector elements

### Using R expressions, update the queue successively for Noah's ark {queue <- c("sheep", "fox", "owl", "ant")}

##### the serpent arrives;

```{r}
queue <- c("sheep", "fox", "owl", "ant") # creating a vector for initial queue of animals to enter into the boat
print(queue)
```

```{r}
q_ser <- c(queue, "serpent") # concatanating serpent to the queue
print(q_ser)
```

##### sheep enters the boat;

```{r}
q_sep <- q_ser[-1] # remove first element from the queue after sheep moves in the boat
print(q_sep)
```

##### donkey arrives and talk his way to the front line;

```{r}
q_donk <- c("Donkey", q_sep) # makes donkey first element by combining it first to existing queue
print(q_donk)
```

##### the serpent gets impatient and leaves;

```{r}
q_ser1 <- q_donk[-5] # serpent leaves by removing 5th element
print(q_ser1)
```

##### owl gets bored and leave;

```{r}
q_owl <- q_ser1[-3] # removing owl 
print(q_owl)
```

##### the aphid arrives and the ant invites him to cut in line;

```{r}

q_aph <- append(x=q_owl, values = "Aphid", after = 2) # merging vector with append function that adds an element to a vector with position assigned
print(q_aph)
```

##### Finally, determine the position of the aphid in the line.

```{r}
which(q_aph == "Aphid") # which position is aphids
```

# Vector of integers

### _Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7_

```{r}
A <- seq(from=1, to=100, by=1) # assigning a variable to list of elements from 1 to 100
print(A)
B <- A [!(!A%%2) + (!A%%3) + (!A%%7)] # creates vector of elements using ! that are not divisible by 2, 3 or 7
print(B)

```