---
title: "Test hypothesis by creating fake datasets"
author: "Ashma"
date: "February 26, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---


## Null Hypothesis 
```
Presence of antimicrobial resistant _Staphylococcus aureus_ strains is not higher in humans than cows in dairy farms.
```
```{r}
# loading packages
library(tidyverse)
library(ggplot2)
library(MASS)
library(dplyr)
```

# Creating a fake data for ANOVA

```{r}
nGroup <- 2
nName <- c("Cows", "Humans")
nSize <- c(25,28) # number of obervations of each group
nMean <- c(50,21) # mean of each group
nSD <- c(10,7) # sd of each group
ID <- 1:(sum(nSize)) # id vector of each rows
AMR <- c(rnorm(n=nSize[1], mean = nMean[1], sd = nSD[1]),
            rnorm(n=nSize[2], mean = nMean[2], sd = nSD[2]))
            
Spp <- rep(nName,nSize) # repeats nName by nSize times
table(Spp)
ano_data <- data.frame(ID,Spp,AMR)
str(ano_data)
```

# Running ANOVA
```{r}
AnoModel <- aov(AMR~Spp, data = ano_data)
summary(AnoModel)
```

# ANOVA plot
```{r}
AnoPlot <- ggplot(data=ano_data,aes(x=Spp,y=AMR,fill=Spp)) +
           geom_boxplot()
print(AnoPlot)
```

```
Looking at the summary results of ANOVA, it shows that there is a significant presence of antibiotic resistance strains in humans than cows.
```
### Altering the means keeping the original sample size to see the variation and significance of data

```{r}
nGroup <- 2
nName <- c("Cows", "Humans")
nSize <- c(25,28) # number of obervations of each group
nMean <- c(12,10) # adjusting means of each group
nSD <- c(10,7) # sd of each group
ID <- 1:(sum(nSize)) # id vector of each rows
AMR <- c(rnorm(n=nSize[1], mean = nMean[1], sd = nSD[1]),
            rnorm(n=nSize[2], mean = nMean[2], sd = nSD[2]))
            
Spp <- rep(nName,nSize) # repeats nName by nSize times
table(Spp)
ano_data <- data.frame(ID,Spp,AMR)
str(ano_data)

AnoModel <- aov(AMR~Spp, data = ano_data)
summary(AnoModel)
AnoPlot <- ggplot(data=ano_data,aes(x=Spp,y=AMR,fill=Spp)) +
           geom_boxplot()
print(AnoPlot)
```

```
Changing means of the groups from 50 to 12 and 21 to 10 made the results insignificant.
```

### Altering the sample size keeping original means to see variation and significance of data

```{r}
nGroup <- 2
nName <- c("Cows", "Humans")
nSize <- c(25,25) # adjusting number of obervations of each group
nMean <- c(50,21) # means of each group
nSD <- c(10,7) # sd of each group
ID <- 1:(sum(nSize)) # id vector of each rows
AMR <- c(rnorm(n=nSize[1], mean = nMean[1], sd = nSD[1]),
            rnorm(n=nSize[2], mean = nMean[2], sd = nSD[2]))
            
Spp <- rep(nName,nSize) # repeats nName by nSize times
table(Spp)
ano_data <- data.frame(ID,Spp,AMR)
str(ano_data)

AnoModel <- aov(AMR~Spp, data = ano_data)
summary(AnoModel)
AnoPlot <- ggplot(data=ano_data,aes(x=Spp,y=AMR,fill=Spp)) +
           geom_boxplot()
print(AnoPlot)
```

```
Keeping same sample size for humans and cows produced significant results and changing the sample size from 25 to 10 still gave significant results.
```
### Conclusion

After adjusting the means and sample size for detecting significant pattern, the sample size and means should be as kept the first code chunk for the results to be significant with p-value less than 0.05.

